---
title: "PCA-example"
author: "Mariachiara Fortuna"
date: "November 4, 2019"
output: html_document
---

```{r}
library(psych)
library(GGally)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE)
```

```{r}
USArrests
```

# Exploratory analysis

Check the correlation:

```{r}
cor(USArrests)
```


Visual representation of correlation between variables:


```{r}
pairs.panels(USArrests)
```

```{r}
ggpairs(USArrests)
```


# PCA analysis

```{r}
pca_arrest <- prcomp(USArrests, center = T, scale = T)
```

```{r}
pca_arrest
```


```{r}
summary(pca_arrest)
```

### Some math

Standard deviation of the components are the square roots of the eigenvalues of the Sigma matrix (variance & cov matrix of X)

These are the eigenvalues (scaling matrix):

```{r}
pca_arrest$sd^2
```


Here we can compute the proportion of variance:

```{r}
pca_arrest$sd^2/sum(pca_arrest$sd^2)
```

These are the eigenvectors (rotation matrix):

```{r}
pca_arrest$rotation
```

Let's do some math:

```{r}
A <- pca_arrest$rotation
D <- diag(pca_arrest$sd^2)
```

```{r}
A %*% D %*% t(A)
```

```{r}
cor(USArrests)
```

Here we can see the new values of the observations in the PC space:

```{r}
pca_arrest$x
```

## Some variations

Select a given number of components:

```{r}
pca_arrest_3 <- prcomp(USArrests, center = T, scale = T,
                     rank. = 3)
```

Select components imposing a tolerance:

```{r}
pca_arrest_tol <- prcomp(USArrests, center = T, scale = T,
                     tol = 0.4)
pca_arrest_tol
```

Doing PCA on a subset of variables:

```{r}
pca_arrest_subset <- prcomp(~ Murder + Assault + Rape, data = USArrests, 
                         center = T, scale = T)
pca_arrest_subset
```


# Bivariate plot


```{r}
pc_data <- as.data.frame(pca_arrest$x[,1:2])
pc_rotation <- as.data.frame(pca_arrest$rotation[,1:2])

```


```{r}
ggplot() +
  geom_point(data = pc_data, aes(x = PC1, y = PC2)) +
  geom_segment(data = pc_rotation, 
               aes(x = 0, y = 0,
                   xend = PC1,
                   yend = PC2),
               arrow = arrow(length=unit(0.10,"cm"),
                             ends="last", type = "closed"),
             col = "red") +
  ggforce::geom_circle(aes(x0 = 0, y0 = 0, r = 1),
                       col = "grey") +
  theme_minimal()
```

```{r}
ggplot(seals, aes(long, lat)) +
  geom_segment(aes(xend = long + delta_long, yend = lat + delta_lat),
    arrow = arrow(length = unit(0.1,"cm"))) +
  borders("state")
```

