---
title: "Preprocessing:Lecture"
author: "Kofi"
date: "07/11/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r cars}
library(tidyverse)
library(dataMaid)
library(caret)
library(psych)
library(GGally)
library(corrplot)

```

```{r}
Data <- readRDS("./data/segmentation.rds")
```

## Preparation the data
```{r}
seg_data <- Data %>%
         filter(Case=='Train') %>%
         select(-c(Cell,Case,Class), -contains("Status"))

```

## Associated Test data

```{r}

seg_data_Test <- Data %>%
         filter(Case=='Test') %>%
         select(-c(Cell,Case,Class), -contains("Status"))
```



## Center and Scale 
```{r}
T_seg <- seg_data %>%
         preProcess(method="center")

seg_centered <-predict(T_seg,seg_data)

```

```{r}
T_scale <- seg_data %>%
         preProcess(method=c("center","scale"))

Seg_scale <- predict(T_scale,seg_data)

```

## Applying PcA

```{r}

T_pca <- seg_data %>%
         preProcess(method=c("center","scale","pca"),
                    thresh = 0.99)

Seg_pca_Train <- predict(T_pca,seg_data)

Seg_pca_Test <-  predict(T_pca,seg_data_Test)


```

### Applying Range on the Train data

```{r}

T_Range <- seg_data %>%
               preProcess(method="range",rangeBounds = c(0, 1))

Seg_Range_train <- predict(T_Range,seg_data )

```

```{r}
 #Hist_data <- seg_data$KurtIntenCh1


Hist_data <- ggplot(seg_data,aes(log(KurtIntenCh1)))+geom_histogram()+
theme_minimal()


Hist_data
```





You can also embed plots, for example:

```{r}
T_boxcox <- seg_data %>%
  preProcess(method='BoxCox')
  
Seg_boxcox <- predict(T_boxcox,seg_data)
```

## Boxcox Singly
```{r}
T_yeojohn <- seg_data %>%
  preProcess(method='YeoJohnson')
  
Seg_yeojohn <- predict(T_yeojohn,seg_data)

```

## Outlier
```{r}

identifyOutliers(seg_data$KurtIntenCh1)
```
It identify the outliers



## Removing Predictors 

```{r}
nearZeroVar(seg_data)

```
No predictor to removed

## Highly Correllated Predictors

```{r}
pairs.panels(seg_data[,1:10])


```
```{r}

ggpairs(Data,columns = 4:7,
        aes(col=Class),progress = FALSE)
```
```{r}
ggduo(data = Data,
      colnames(Data[,4:7]),
      "Class")
```
```{r}
Seg_heat <- cor(seg_data)

corrplot::corrplot(Seg_heat,
                     order="hclust",
                     tl.cex=0.3)


```
```{r}
hicor<- findCorrelation(Seg_heat, cutoff = 0.80)

SegFilter <- seg_data[,-hicor]

```
##  Exercise Preprocess

```{r}
Glass <- readRDS("./data/glass.rds")
dim(Glass)

```
## Explore Relationship Between Predictors

```{r}
Glass_dat <- Glass[,-10]

pairs.panels(Glass_dat)



```

   We know from the  above plot that there are many most of the predictor are negatively correlated with each other. With few of the element been correlated to each other in a positive manner. The highest positive correlation occured between the Calcium and the refractive index. 





##  Question 2a:Checking For Outliers Within The Data

```{r}
boxplot(Glass_dat)




```

The above is a strong indicator which is explain which predictors have a outliers. In the above plot can see that Refractive index,magnissium and Iron have **minimium** outliers. The remaining predictors on the global scale all have outliers.


## Question 2b: Checking For Skewness Within The Predictors

```{r}
T_boxcox <- seg_data %>%
  preProcess(method='BoxCox')
  
Seg_boxcox <- predict(T_boxcox,seg_data)



```

